Component response rate variation drives stability in large complex systems

### 2018-06-18 notebook/arxiv/ms.tex
Change the wording in the SI
Add updated (and corrected) table to SI
Update the simulation number total

To further investigate the potential of \(Var(\gamma)\) to be stabilising,
I used a genetic algorithm (the space of possible \(\gamma\)
values was too large to search
exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{18}};
see \hyperlink{SIstart}{Supplementary Information}). For each of 10000 
random \(M\), the
genetic algorithm initialised 1000 different sets of
\(\gamma \sim \mathcal{U}(0, 2)\) values of size \(S\). Eigenanalysis was
performed on \(M\) using 
each set of \(\gamma\) values, and the 20 sets with the
lowest \(\max\left(\Re(\lambda)\right)\) each produced 50 clonal offspring with
subsequent mutation and crossover between the resulting new population
of 1000 \(\gamma\) sets. The genetic algorithm terminated if a stable
\(M\) was found, 20 generations occurred, or a convergence criteria of
minimum fitness increase between generations was satisfied. Across
\(S = \{2, 3, ..., 39, 40\}\), sets of \(\gamma\) values were found that
resulted in stable systems with probabilities that were up to four orders of
magnitude higher than when \(\gamma = 1\) (see \hyperlink{SIstart}{Supplementary
Information}), meaning that stability could often be achieved by 
manipulating \(S\) \(\gamma\) values rather than \(S \times S\) \(M\) elements.
Hence, managing the response rates of system components in a targetted
way can potentially facilitate the stabilisation of complex systems 
through a reduction in dimensionality.

@@

Some stability results from this table are shown below. Results for \texttt{A0}
indicate systems in which \(\gamma = 1\), while \texttt{A1} refers to systems in 
which the genetic algorithm searched for a set of \(\gamma\) values that 
stabilised the system.

\begin{longtable}[]{@{}rrrrrrr@{}}
\toprule
N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
A1\_stabilised & A1\_destabilised\tabularnewline
\midrule
\endhead
2 & 4 & 9996 & 4 & 9996 & 0 & 0\tabularnewline
3 & 42 & 9958 & 42 & 9958 & 0 & 0\tabularnewline
4 & 133 & 9867 & 133 & 9867 & 0 & 0\tabularnewline
5 & 414 & 9586 & 411 & 9589 & 3 & 0\tabularnewline
6 & 809 & 9191 & 799 & 9201 & 10 & 0\tabularnewline
7 & 1380 & 8620 & 1339 & 8661 & 41 & 0\tabularnewline
8 & 2074 & 7926 & 1927 & 8073 & 147 & 0\tabularnewline
9 & 2885 & 7115 & 2503 & 7497 & 382 & 0\tabularnewline
10 & 3842 & 6158 & 3158 & 6842 & 684 & 0\tabularnewline
11 & 4867 & 5133 & 3613 & 6387 & 1255 & 1\tabularnewline
12 & 5932 & 4068 & 4148 & 5852 & 1784 & 0\tabularnewline
13 & 6937 & 3063 & 4470 & 5530 & 2468 & 1\tabularnewline
14 & 7784 & 2216 & 4724 & 5276 & 3060 & 0\tabularnewline
15 & 8519 & 1481 & 5086 & 4914 & 3433 & 0\tabularnewline
16 & 9081 & 919 & 5262 & 4738 & 3819 & 0\tabularnewline
17 & 9431 & 569 & 5368 & 4632 & 4063 & 0\tabularnewline
18 & 9671 & 329 & 5571 & 4429 & 4100 & 0\tabularnewline
19 & 9844 & 156 & 5807 & 4193 & 4037 & 0\tabularnewline
20 & 9934 & 66 & 6133 & 3867 & 3801 & 0\tabularnewline
21 & 6387 & 34 & 6421 & 3579 & 3545 & 0\tabularnewline
22 & 6634 & 11 & 6645 & 3355 & 3344 & 0\tabularnewline
23 & 7037 & 8 & 7045 & 2955 & 2947 & 0\tabularnewline
24 & 7468 & 3 & 7471 & 2529 & 2526 & 0\tabularnewline
25 & 7816 & 0 & 7816 & 2184 & 2184 & 0\tabularnewline
26 & 8192 & 0 & 8192 & 1808 & 1808 & 0\tabularnewline
27 & 8680 & 0 & 8680 & 1320 & 1320 & 0\tabularnewline
28 & 8936 & 0 & 8936 & 1064 & 1064 & 0\tabularnewline
29 & 9296 & 0 & 9296 & 704 & 704 & 0\tabularnewline
30 & 9523 & 0 & 9523 & 477 & 477 & 0\tabularnewline
31 & 9705 & 0 & 9705 & 295 & 295 & 0\tabularnewline
32 & 9816 & 0 & 9816 & 184 & 184 & 0\tabularnewline
33 & 9894 & 0 & 9894 & 106 & 106 & 0\tabularnewline
34 & 9941 & 0 & 9941 & 59 & 59 & 0\tabularnewline
35 & 9968 & 0 & 9968 & 32 & 32 & 0\tabularnewline
36 & 9991 & 0 & 9991 & 9 & 9 & 0\tabularnewline
37 & 9993 & 0 & 9993 & 7 & 7 & 0\tabularnewline
38 & 9999 & 0 & 9999 & 1 & 1 & 0\tabularnewline
39 & 9999 & 0 & 9999 & 1 & 1 & 0\tabularnewline
40 & 10000 & 0 & 10000 & 0 & 0 & 0\tabularnewline
\bottomrule
\end{longtable}

>>>>

To further investigate the potential of \(Var(\gamma)\) to be stabilising,
I used a genetic algorithm (the space of possible \(\gamma\)
values was too large to search
exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{18}};
see \hyperlink{SIstart}{Supplementary Information}). For each of 40000 
random \(M\), the
genetic algorithm initialised 1000 different sets of
\(\gamma \sim \mathcal{U}(0, 2)\) values of size \(S\). Eigenanalysis was
performed on \(M\) using 
each set of \(\gamma\) values, and the 20 sets with the
lowest \(\max\left(\Re(\lambda)\right)\) each produced 50 clonal offspring with
subsequent mutation and crossover between the resulting new population
of 1000 \(\gamma\) sets. The genetic algorithm terminated if a stable
\(M\) was found, 20 generations occurred, or a convergence criteria of
minimum fitness increase between generations was satisfied. Across
\(S = \{2, 3, ..., 39, 40\}\), sets of \(\gamma\) values were found that
resulted in stable systems with probabilities that were up to four orders of
magnitude higher than when \(\gamma = 1\) (see \hyperlink{SIstart}{Supplementary
Information}), meaning that stability could often be achieved by 
manipulating \(S\) \(\gamma\) values rather than \(S \times S\) \(M\) elements.
Hence, managing the response rates of system components in a targetted
way can potentially facilitate the stabilisation of complex systems 
through a reduction in dimensionality.

@@

Stability results for 40000 \(M\) for each \(S\) from 2-40 are shown below. Results for \texttt{A0}
indicate systems in which \(\gamma = 1\), while \texttt{A1} refers to systems in 
which the genetic algorithm searched for a set of \(\gamma\) values that 
stabilised the system.

\begin{longtable}[]{@{}rrrrrrr@{}}
\toprule
S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
A1\_stabilised & A1\_destabilised\tabularnewline
\midrule
\endhead
2 & 13 & 39987 & 13 & 39987 & 0 & 0\tabularnewline
3 & 157 & 39843 & 157 & 39843 & 0 & 0\tabularnewline
4 & 563 & 39437 & 563 & 39437 & 0 & 0\tabularnewline
5 & 1588 & 38412 & 1581 & 38419 & 7 & 0\tabularnewline
6 & 3268 & 36732 & 3219 & 36781 & 49 & 0\tabularnewline
7 & 5370 & 34630 & 5196 & 34804 & 175 & 1\tabularnewline
8 & 8231 & 31769 & 7604 & 32396 & 627 & 0\tabularnewline
9 & 11704 & 28296 & 10181 & 29819 & 1523 & 0\tabularnewline
10 & 15215 & 24785 & 12303 & 27697 & 2912 & 0\tabularnewline
11 & 19529 & 20471 & 14599 & 25401 & 4930 & 0\tabularnewline
12 & 23480 & 16520 & 16210 & 23790 & 7270 & 0\tabularnewline
13 & 27570 & 12430 & 17800 & 22200 & 9770 & 0\tabularnewline
14 & 31096 & 8904 & 18899 & 21101 & 12198 & 1\tabularnewline
15 & 34036 & 5964 & 19948 & 20052 & 14089 & 1\tabularnewline
16 & 36231 & 3769 & 21000 & 19000 & 15231 & 0\tabularnewline
17 & 37737 & 2263 & 21693 & 18307 & 16044 & 0\tabularnewline
18 & 38808 & 1192 & 22291 & 17709 & 16518 & 1\tabularnewline
19 & 39391 & 609 & 23221 & 16779 & 16170 & 0\tabularnewline
20 & 39711 & 289 & 23943 & 16057 & 15768 & 0\tabularnewline
21 & 39867 & 133 & 25361 & 14639 & 14506 & 0\tabularnewline
22 & 39950 & 50 & 26584 & 13416 & 13366 & 0\tabularnewline
23 & 39982 & 18 & 28162 & 11838 & 11820 & 0\tabularnewline
24 & 39994 & 6 & 29660 & 10340 & 10334 & 0\tabularnewline
25 & 39997 & 3 & 31405 & 8595 & 8592 & 0\tabularnewline
26 & 39999 & 1 & 33057 & 6943 & 6942 & 0\tabularnewline
27 & 40000 & 0 & 34602 & 5398 & 5398 & 0\tabularnewline
28 & 40000 & 0 & 35957 & 4043 & 4043 & 0\tabularnewline
29 & 40000 & 0 & 37154 & 2846 & 2846 & 0\tabularnewline
30 & 40000 & 0 & 38104 & 1896 & 1896 & 0\tabularnewline
31 & 40000 & 0 & 38722 & 1278 & 1278 & 0\tabularnewline
32 & 40000 & 0 & 39259 & 741 & 741 & 0\tabularnewline
33 & 40000 & 0 & 39551 & 449 & 449 & 0\tabularnewline
34 & 40000 & 0 & 39742 & 258 & 258 & 0\tabularnewline
35 & 40000 & 0 & 39879 & 121 & 121 & 0\tabularnewline
36 & 40000 & 0 & 39929 & 71 & 71 & 0\tabularnewline
37 & 40000 & 0 & 39967 & 33 & 33 & 0\tabularnewline
38 & 40000 & 0 & 39989 & 11 & 11 & 0\tabularnewline
39 & 40000 & 0 & 39994 & 6 & 6 & 0\tabularnewline
40 & 40000 & 0 & 39996 & 4 & 4 & 0\tabularnewline
\bottomrule
\end{longtable}
